### 2025-04-25 NHSH_hospitalizations_seed/ES_ARIMA/ES_ARIMA.Rmd
update

---
title: "ARIMA ENSEMBLE MODELS - Influenza hospitalization data"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMA models. The function fits a rolling window of 104 weeks to generate forecasts. It returns some metrics that evaluate the performance of the models: target_end_date, abs_error, cases, forecast, number of models, weighted interval score (WIS) and predictive quantiles (%). The user defines if it will use an AUTO ARIMA (auto=TRUE), or ensembles of 27 (ES27=TRUE) or ensembles of 64 models (ES64=TRUE). It can also choose the number of weeks ahead for each forecast, and the size of the rolling window which is set as 104 (2 years).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!!
LOADING THE PACKAGES
!!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the ADJACENT states models
source("ES_ARIMA.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)

```

AUTO ARIMA WEEK1

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK1_list <- mclapply(list_of_states, ES_ARIMA, auto=TRUE, n_weeks_ahead=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK1 <- bind_rows(AUTO_ARIMA_WEEK1_list, .id = "State")
```

AUTO ARIMA WEEK2

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK2_list <- mclapply(list_of_states , ES_ARIMA, auto=TRUE, n_weeks_ahead=2, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK2 <- bind_rows(AUTO_ARIMA_WEEK2_list, .id = "State")

```

AUTO ARIMA WEEK3

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK3_list <- mclapply(list_of_states , ES_ARIMA, auto=TRUE, n_weeks_ahead=3, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK3 <- bind_rows(AUTO_ARIMA_WEEK3_list, .id = "State")

```

AUTO ARIMA WEEK4

```{r}
start_time <- Sys.time()

AUTO_ARIMA_WEEK4_list <- mclapply(list_of_states, ES_ARIMA, auto=TRUE, n_weeks_ahead=4 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_ARIMA_WEEK4 <- bind_rows(AUTO_ARIMA_WEEK4_list, .id = "State")

```

```{r}
save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
```

ES27 ARIMA WEEK1

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK1_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK1 <- bind_rows(ES27_ARIMA_WEEK1_list, .id = "State")

```

ES27 ARIMA WEEK2

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK2_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=2, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK2 <- bind_rows(ES27_ARIMA_WEEK2_list, .id = "State")

```

ES27 ARIMA WEEK3

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK3_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=3, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK3 <- bind_rows(ES27_ARIMA_WEEK3_list, .id = "State")

```

ES27 ARIMA WEEK4

```{r}
start_time <- Sys.time()

ES27_ARIMA_WEEK4_list <- mclapply(list_of_states, ES_ARIMA, ES27=TRUE, n_weeks_ahead=4 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_ARIMA_WEEK4 <- bind_rows(ES27_ARIMA_WEEK4_list, .id = "State")

```

```{r}
save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
```

ES64 ARIMA WEEK1

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK1_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK1 <- bind_rows(ES64_ARIMA_WEEK1_list, .id = "State")

```

ES64 ARIMA WEEK2

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK2_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=2, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK2 <- bind_rows(ES64_ARIMA_WEEK2_list, .id = "State")

```

ES64 ARIMA WEEK3

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK3_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=3, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK3 <- bind_rows(ES64_ARIMA_WEEK3_list, .id = "State")

```

ES64 ARIMA WEEK4

```{r}
start_time <- Sys.time()

ES64_ARIMA_WEEK4_list <- mclapply(list_of_states, ES_ARIMA, ES64=TRUE, n_weeks_ahead=4 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_ARIMA_WEEK4 <- bind_rows(ES64_ARIMA_WEEK4_list, .id = "State")

```

```{r}
save.image("ARIMA_MODELS_influenza_hospitalization.Rdata")
```