### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/ES_EPIWEEK/EPIWEEK.Rmd

---
title: "EPIWEEK ARIMAX MODELS - Influenza hospitalization data"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

This function utilizes ensembles and single automatic ARIMAX models which have mean cases by epidemiological weeks over the last 2 years as exogenous variables. The function fits on a rolling window of previous 104 weeks for the state under analysis and a rolling windows with the same size with 1 week-lag for the exogenous variables to generate forecasts. It return some metrics that evaluate the performance of the models:target_end_date, abs_error, cases, forecast, 'N_of_models", weighted interval score (WIS), predictive quantiles. The user can choose a  single best automatic ARIMAXs (auto=TRUE), or ensembles of 27 permutations of 0,1,2 pdq's (ES27=TRUE) or 64 permutations of 0,1,2,3 pdq's (ES64=TRUE). The user also chooses the number of weeks ahead for each forecast, and the size of the rolling window which is set as 2 years (104 weeks).

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
!!!!!!!!!!!!!!!!!!!!
LOADING THE PACKAGES
!!!!!!!!!!!!!!!!!!!!
```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
```
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
LOADING DATASET AND FUNCTIONS 
!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
```{r}
###########################################
#       LOADING AND CLEANING THE DATASET  #        
###########################################

# Loads the ADJACENT states models
source("ES_EPIWEEK_nolog.R", local = TRUE, chdir = TRUE)

# Loads the ILI dataset
my_data = read.csv("treated_influenza_hosp_dataframe_v2.csv")
my_data$target_end_date<-as.Date(my_data$target_end_date) # set the dates as dates

list_of_states <- split(my_data, my_data$state_name)
```

AUTO EPIWEEK WEEK1

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_EPIWEEK_WEEK1_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=1,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_EPIWEEK_WEEK1 <- bind_rows(AUTO_EPIWEEK_WEEK1_list, .id = "State")

# SAVE TO CORRECT FOLDER
#write.csv(EPIWEEK/AUTO/AUTO_EPIWEEK_WEEK1, file = "AUTO_EPIWEEK_WEEK1.csv", row.names = TRUE)
```

AUTO EPIWEEK WEEK2

```{r}
start_time <- Sys.time()

# RUN MODEL 
AUTO_EPIWEEK_WEEK2_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=2,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# FINAL DATAFRAME
AUTO_EPIWEEK_WEEK2 <- bind_rows(AUTO_EPIWEEK_WEEK2_list, .id = "State")

```

AUTO EPIWEEK WEEK3

```{r}
start_time <- Sys.time()

AUTO_EPIWEEK_WEEK3_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=3,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_EPIWEEK_WEEK3 <- bind_rows(AUTO_EPIWEEK_WEEK3_list, .id = "State")

```

AUTO EPIWEEK WEEK4

```{r}
start_time <- Sys.time()

AUTO_EPIWEEK_WEEK4_list <- mclapply(list_of_states, ES_EPIWEEK, auto=TRUE, n_weeks_ahead=4,week_lag=1 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
AUTO_EPIWEEK_WEEK4 <- bind_rows(AUTO_EPIWEEK_WEEK4_list, .id = "State")

```

```{r}
save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")

```

ES27 EPIWEEK WEEK1

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK1_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=1,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK1 <- bind_rows(ES27_EPIWEEK_WEEK1_list, .id = "State")

```

ES27 EPIWEEK WEEK2

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK2_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=2,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK2 <- bind_rows(ES27_EPIWEEK_WEEK2_list, .id = "State")

```

ES27 EPIWEEK WEEK3

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK3_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=3,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK3 <- bind_rows(ES27_EPIWEEK_WEEK3_list, .id = "State")

```

ES27 EPIWEEK WEEK4

```{r}
start_time <- Sys.time()

ES27_EPIWEEK_WEEK4_list <- mclapply(list_of_states, ES_EPIWEEK, ES27=TRUE, n_weeks_ahead=4,week_lag=1 ,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES27_EPIWEEK_WEEK4 <- bind_rows(ES27_EPIWEEK_WEEK4_list, .id = "State")

```

```{r}
save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")

```

ES64 EPIWEEK WEEK1

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK1_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=1,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK1 <- bind_rows(ES64_EPIWEEK_WEEK1_list, .id = "State")

```

ES64 EPIWEEK WEEK2

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK2_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=2,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK2 <- bind_rows(ES64_EPIWEEK_WEEK2_list, .id = "State")

```

ES64 EPIWEEK WEEK3

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK3_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=3,week_lag=1, mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK3 <- bind_rows(ES64_EPIWEEK_WEEK3_list, .id = "State")

```

ES64 EPIWEEK WEEK4

```{r}
start_time <- Sys.time()

ES64_EPIWEEK_WEEK4_list <- mclapply(list_of_states, ES_EPIWEEK, ES64=TRUE, n_weeks_ahead=4,week_lag=1,mc.cores=2)%>%
  setNames(names(list_of_states))

end_time <- Sys.time()
run_time <- end_time - start_time

print(run_time)

# Combine the list of data frames into a single data frame with names as a column
ES64_EPIWEEK_WEEK4 <- bind_rows(ES64_EPIWEEK_WEEK4_list, .id = "State")

```

```{r}
save.image("EPIWEEK_MODELS_influenza_hospitalization_nolog.Rdata")

```