### 2025-04-25 NHSH_hospitalizations_seed_nolog_correct/flu_hospitalization_dataset .Rmd

---
title: "Influenza Hopistalizations dataset with Florida - NHSN"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading datasets

```{r}
library(purrr)
library(tidyr)
library(dplyr)
library(lubridate)
library(imputeTS)
library(MMWRweek)
```

Influenza hospitalizations

```{r}
influenza_data<-read.csv("Weekly_United_States_Hospitalization_Metrics_by_Jurisdiction__During_Mandatory_Reporting_Period_from_August_1__2020_to_April_30__2024__and_for_Data_Reported_Voluntarily_Beginning_May_1__2024__National_Healthcare_Safety_Network__NHSN__20241001.csv")

state_codes<-read.csv("State_Codes.csv")

```

Choosing U.S. states

```{r}
states=state_codes$abbreviation[1:48] # removing Florida
```

Creating a list with multiple dataframes

```{r}

influenza_data$Week.Ending.Date<-as.Date(influenza_data$Week.Ending.Date)
influenza_data$epiweek<-MMWRweek(influenza_data$Week.Ending.Date)$MMWRweek

# Initialize an empty list to store the results
influenza_hosp_by_state <- list()

# Loop through each state in the 'states' vector
for (i in 1:48) {
  st <- states[i]  # Get the current state
  # Filter and arrange the data for the current state
  influenza_hosp_by_state[[i]] <- influenza_data %>%
    arrange(Week.Ending.Date) %>%  # Order by date
    filter(Geographic.aggregation == st)  # Filter for the current state
}

```

Finding the datasets that have NAs

We need to start from week 12 of all datasets till we remove all initial NAs.
In addition to the initial week 3 states had more NAs on three states, which we replace with zeros.

WV(state code=45): 3 NAs
MN(state code=20): 18 NAs
MA(state code=18): 32 NAs

```{r}
#INITIAL NAs
# Loop through each element in state_data_list
for (i in 1:length(influenza_hosp_by_state)) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  print(sum(is.na(influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:18])))
}

#MORE NAs
# Loop through each element in state_data_list
for (i in 1:48) {

  state_name <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  missing_values_count <- sum(is.na(influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:length(influenza_hosp_by_state[[i]]$Week.Ending.Date)]))

  print(paste("State:", state_name,"State code:", i, "- Missing values:", missing_values_count))
}

#FINAL NAs
# Loop through each element in state_data_list
for (i in 1:length(influenza_hosp_by_state)) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  print(sum(is.na(influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[190:198])))
}

```

3 states with NAs that will be replaced with zeros.

```{r}

# Loop through each element in state_data_list
for (i in c(18,20,45)) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  plot(influenza_hosp_by_state[[i]]$Week.Ending.Date[12:198],influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198], main = states_names, ylab="Cases", xlab="Dates")
}
```
Noting that the weeks that have NAs are at the end of the last season and in a location with low influenza activity we will replace these values with zero.

```{r}
for (i in c(18, 20, 45)) {
  # Extract the unique state name from the 'Geographic.aggregation' column
  state_name <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)

  # Extract the 'Weekly.Total.Influenza.Admissions' column from 12 to 198
  admissions <- influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198]

  # Replace only the NA values with zero
  admissions[is.na(admissions)] <- 0

  # Update the original data with the new values
  influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198] <- admissions

  # Plot the time series and set the title to the state name
  plot(admissions, 
       type = "l",   # Line plot
       main = state_name, 
       xlab = "Week", 
       ylab = "Influenza Admissions")
}

```
The clean final dataset for all states

```{r}
# Loop through each element in state_data_list
for (i in 1:48) {

  # Extract the unique state name from the 'state' column
  states_names <- unique(influenza_hosp_by_state[[i]]$Geographic.aggregation)
  # Plot epi_week, setting the title to the state name
  plot(influenza_hosp_by_state[[i]]$Week.Ending.Date[12:198],influenza_hosp_by_state[[i]]$Weekly.Total.Influenza.Admissions[12:198], main = states_names)
}

# Combine the list of data frames into a single data frame with names as a column
influenza_hosp_dataframe <- bind_rows(lapply(influenza_hosp_by_state, function(df) df[12:198, ]))
```

Final dataset similar in format that can be read by the ensembles

```{r}
colnames(state_codes)[3] <- "state_name"

final_dataset <- influenza_hosp_dataframe %>%
  select("target_end_date"=Week.Ending.Date, "abbreviation"=Geographic.aggregation,"cases"= Weekly.Total.Influenza.Admissions)

final_dataset <- left_join(final_dataset, state_codes, by = "abbreviation")
final_dataset

# write the dataset to use in the models
write.csv(final_dataset, file = "ES_ADJACENT/treated_influenza_hosp_dataframe_v2.csv")
write.csv(final_dataset, file = "ES_ARIMA/treated_influenza_hosp_dataframe_v2.csv")
write.csv(final_dataset, file = "ES_EPIWEEK/treated_influenza_hosp_dataframe_v2.csv")
write.csv(final_dataset, file = "ES_TEMPERATURE/treated_influenza_hosp_dataframe_v2.csv")

```