### 2025-05-07 ERA5 data/2010-2024/ERA5_2010_2024.Rmd

---
title: "ERA5 Temperature and Humidity weekly data 2010-2024"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document converts ERA5 NetCDF's with daily data into weekly data for each state on the contiguous U.S.

```{r}
library(terra)
library(dplyr)
library(lubridate)
library(tidyr)
library(parallel)
library(ncdf4)
library(MMWRweek)
```

Downloading temperature and humidity data from the cloud.

```{python}
# Importando a biblioteca do xarray e outras auxiliares
import xarray as xr
import numpy as np
import os
import matplotlib.pyplot as plt
from mpl_toolkits.axes_grid1.inset_locator import inset_axes
import datetime as dt
import pandas as pd
import geopandas
import regionmask
import seaborn as sns
import pytz
import warnings
warnings.filterwarnings("ignore")

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
temperature = xr.open_dataset(url+'Air_Temperature')
temperature =temperature.sel(time=slice("2010-01-01","2024-07-12"))
temperature = temperature.assign_coords(lon=(((temperature.lon + 180) % 360) - 180))
temperature
# xarray dataset
temperature=temperature.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
temperature.to_netcdf(path='ERA5_temperature_data_US_2010_2024.nc')

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
humidity= xr.open_dataset(url+'Specific_humidity')
humidity=humidity.sel(time=slice("2010-01-01","2024-07-12"))
humidity= humidity.assign_coords(lon=(((humidity.lon + 180) % 360) - 180))
humidity=humidity.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
# xarray dataset
humidity.to_netcdf(path='ERA5_humidity_data_US_2010_2024.nc')
```

Plotting temperature and humidity NetCDFs.

```{r}
temp <- rast('ERA5_temperature_US_2010_2024.nc')  # Temperature
humi <- rast('ERA5_humidity_data_US_2010_2024.nc')  # Humidity

# Calculate the mean over the time dimension
temp_mean <- mean(temp, na.rm = TRUE)

# Calculate the mean over the time dimension
humi_mean <- mean(humi, na.rm = TRUE)

# Plot
plot(temp_mean)
plot(humi_mean)
```

Here I will load and inspect the states' shapefiles. I will use them later in order to calculate weekly temperature and humidity for each state.

```{r}
# Codes and abbreviations of each U.S. state
state_code<-read.csv("State_Codes.csv")

# Read the shapefile
US <- vect("states_shapefile/States_shapefile.shp")

plot(US)

# Loop through the first 48 state codes
for (i in 1:48) {
    code <- state_code$abbreviation[i]

    # Filter the SpatVector based on the state code
    state_shapefile <- US[US$State_Code == code, ]

    # Plot the filtered state shapefile
    plot(state_shapefile, main = paste("State:", code))
}

state_code
```

Here I develop a function that take states codes for shapefiles selection and NetCDFs with daily resolution (temperature or humidity) as inputs. It then creates a dataframe with weekly values for each U.S. state that will be utilized as covariates.

```{r}
ERA5_weekly_dataframe<-function(state_code=state_code, my_netcdf) {
  # Filter the state shapefile for the current state
  state_shapefile <- US[US$State_Code == state_code$abbreviation, ]
  # Create a mask for the state using terra's mask function
  netcdf_state_daily <- mask(my_netcdf, state_shapefile)
  # Extract the time dimension (dates)
  dates <- time(netcdf_state_daily)
  # Calculate the mean for each day across the spatial dimensions (lat, lon)
  daily_state_mean_ts <- global(netcdf_state_daily, mean, na.rm = TRUE)
  # Convert the daily time series to a data frame
  daily_state_df <- as.data.frame(daily_state_mean_ts, xy = FALSE, na.rm = TRUE)
  # Add the time column to the data frame
  daily_state_df$time <- as.Date(dates)
  # Calculate the epidemiological week and include as a column
  daily_state_df$epi_week <- MMWRweek(daily_state_df$time)$MMWRweek
  # Get the year and include as a column
  daily_state_df$year <- MMWRweek(daily_state_df$time)$MMWRyear
  # Calculate the mean temperature for each epidemiological week and year
  weekly_mean_value_by_state <- daily_state_df %>%
    group_by(year, epi_week) %>%
    summarise(weekly_value = mean(mean, na.rm = TRUE), .groups = "drop") 
  # Rename the column to the state's location name
  colnames(weekly_mean_value_by_state)[3] <- state_code$location_name
  # Return the weekly data for the current state
  return(weekly_mean_value_by_state)
}
```

Here I select only the abbreviations and state codes of states on the contiguous U.S.

```{r}
# Split dataframe into a list of dataframes by row
states_codes <- split(state_code, seq(nrow(state_code)))
states_codes<-states_codes[1:48] # select only states up to Florida
```

Here I run the function to calculate mean weekly TEMPERATURE for all states on the contiguous U.S.

```{r}
# Run using mclapply

temperature_list_of_dataframes <- mclapply(states_codes, function(state_code) {
  ERA5_weekly_dataframe(state_code = state_code, my_netcdf = temp)
}, mc.cores = 1)

final_temperature_dataframe <- bind_rows(temperature_list_of_dataframes) %>%
   mutate(date = MMWRweek2Date(MMWRyear = year, MMWRweek = epi_week)+6)%>% # I need to add 6 to get the end_week_day
   group_by(date) %>%
   summarize(across(everything(), mean, na.rm = TRUE))

final_temperature_dataframe

write.csv(final_temperature_dataframe, file = "final_temperature_data_2010_2024.csv")
```

Here I run the function to calculate mean weekly SPECIFIC HUMIDITY for all states on the contiguous U.S.

```{r}
# Run using mclapply
humidity_df <- mclapply(states_codes, function(state_code) {
  ERA5_weekly_dataframe(state_code = state_code, my_netcdf = humi)
}, mc.cores = 1)

final_humidity_dataframe <- bind_rows(humidity_df) %>%
   mutate(date = MMWRweek2Date(MMWRyear = year, MMWRweek = epi_week)+6)%>%
   group_by(date) %>%
   summarize(across(everything(), mean, na.rm = TRUE))

write.csv(final_humidity_dataframe , file = "final_humidity_data_2010_2024.csv")
```