### 2025-06-12 ERA5 data/1990-2020/ERA5_1990_2020.Rmd
update

---
title: "ERA5 Temperature and Humidity climatologies for regression analysis"
author: "Victor Felix"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document converts ERA5 NetCDF's with daily data into weekly data for each state on the contiguous U.S.

```{r}
library(terra)
library(dplyr)
library(lubridate)
library(tidyr)
library(parallel)
library(ncdf4)
library(MMWRweek)
```

Downloading temperature and humidity data from the cloud.
You may run this code in python.

```{python}
# Libraries
import xarray as xr
import numpy as np
import os
import matplotlib.pyplot as plt
from mpl_toolkits.axes_grid1.inset_locator import inset_axes
import datetime as dt
import pandas as pd
import geopandas
import regionmask
import seaborn as sns
import pytz
import warnings
warnings.filterwarnings("ignore")

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
temperature = xr.open_dataset(url+'Air_Temperature')
temperature =temperature.sel(time=slice("1990-01-01","2020-12-31"))
temperature = temperature.assign_coords(lon=(((temperature.lon + 180) % 360) - 180))
temperature
# xarray dataset
temperature=temperature.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
temperature.to_netcdf(path='ERA5_temperature_data_US_2010_2024.nc')

url='http://apdrc.soest.hawaii.edu/dods/public_data/Reanalysis_Data/ERA5/daily_3d/'
humidity= xr.open_dataset(url+'Specific_humidity')
humidity=humidity.sel(time=slice("1990-01-01","2020-12-31"))
humidity= humidity.assign_coords(lon=(((humidity.lon + 180) % 360) - 180))
humidity=humidity.sel(lev=1000, lat=slice(20,60),lon=slice(-130,-60))
# xarray dataset
humidity.to_netcdf(path='ERA5_humidity_data_US_2010_2024.nc')
```

After downloading the file you can use the code below to get mean weekly temperature for each state.

Plotting temperature and humidity NetCDFs.

```{r}
temp <- rast('ERA5_temperature_US_1990_2020.nc')  # Temperature
humi <- rast('ERA5_humidity_data_US_1990_2020.nc')  # Temperature

# Calculate the mean over the time dimension
temp_mean <- mean(temp, na.rm = TRUE)

# Calculate the mean over the time dimension
humi_mean <- mean(humi, na.rm = TRUE)

# Plot
plot(temp_mean)
plot(humi_mean)
```

Here I will load and inspect the states shapefiles. I will use them later in order to calculate weekly temperature and humidity for each state.

```{r}
# Codes and abbreviations of each U.S. state
state_code<-read.csv("State_Codes.csv")

# Read the shapefile
US <- vect("states_shapefile/States_shapefile.shp")

plot(US)

# Loop through the first 48 state codes
for (i in 1:48) {
    code <- state_code$abbreviation[i]

    # Filter the SpatVector based on the state code
    state_shapefile <- US[US$State_Code == code, ]

    # Plot the filtered state shapefile
    plot(state_shapefile, main = paste("State:", code))
}

state_code
```

Here I develop a function that take states codes (for shapefiles selection) and NetCDFs with daily resolution (temperature or humidity) as inputs, and creates a dataframe with weekly values for each U.S. state.

```{r}
ERA5_mean_by_state<-function(state_code=state_code, my_netcdf) {
  # Filter the state shapefile for the current state
  state_shapefile <- US[US$State_Code == state_code$abbreviation, ]
  # Create a mask for the state using terra's mask function
  netcdf_state_daily <- mask(my_netcdf, state_shapefile)
  # Calculate the mean for a given state
  temp_mean <- mean(netcdf_state_daily, na.rm = TRUE)  
  # Calculate the mean across the spatial dimensions (lat, lon)
  state_mean <- global(temp_mean, mean, na.rm = TRUE)
    # Rename the column to the state's location name
  colnames(state_mean) <- state_code$location_name
  return(state_mean)
}
```

Here I select only the abbreviations and state codes of states on the contiguous U.S.

```{r}
# Split dataframe into a list of dataframes by row
states_codes <- split(state_code, seq(nrow(state_code)))
states_codes<-states_codes[1:48] # select only states up to Florida

```

Here I run the function to calculate mean weekly TEMPERATURE for all states on the contiguous U.S.

```{r}
# Run using mclapply

mean_temperature_each_state_1990_2020 <- mclapply(states_codes, function(state_code) {
  ERA5_mean_by_state(state_code = state_code, my_netcdf = temp)
}, mc.cores = 1)

mean_temperature_each_state_1990_2020_df<-bind_cols(mean_temperature_each_state_1990_2020)
mean_temperature_each_state_1990_2020_df<-data.frame(t(mean_temperature_each_state_1990_2020_df))
mean_temperature_each_state_1990_2020_df$STATE <-rownames(mean_temperature_each_state_1990_2020_df)
rownames(mean_temperature_each_state_1990_2020_df) <- NULL
mean_temperature_each_state_1990_2020_df
write.csv(mean_temperature_each_state_1990_2020_df, file = "temperature_climatology_1990_2020.csv")
```

Here I run the function to calculate mean weekly SPECIFIC HUMIDITY for all states on the contiguous U.S.

```{r}
mean_humidity_each_state_1990_2020 <- mclapply(states_codes, function(state_code) {
  ERA5_mean_by_state(state_code = state_code, my_netcdf = humi)
}, mc.cores = 1)

mean_humidity_each_state_1990_2020_df<-bind_cols(mean_humidity_each_state_1990_2020)
mean_humidity_each_state_1990_2020_df<-data.frame(t(mean_humidity_each_state_1990_2020_df))
mean_humidity_each_state_1990_2020_df$STATE <-rownames(mean_humidity_each_state_1990_2020_df)
rownames(mean_humidity_each_state_1990_2020_df) <- NULL
mean_humidity_each_state_1990_2020_df
write.csv(mean_humidity_each_state_1990_2020_df, file = "humidity_climatology_1990_2020.csv")
```